[Re] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model

### 2018-03-02 article/ShimouraR-KamijiNL-PenaRFO-CordeiroVL-CeballosCC-RomaroC-RoqueAC-2017.tex
Uploaded version of the manuscript

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C. Potjans,  M. Diesmann, Cerebral Cortex, 24 (3): 785-806, 2014. DOI: 10.1093/cercor/bhs358
  \end{itemize}\par
\end{flushleft}
\end{rebox}

@@

Additionally to these inputs, in Fig.~\ref{fig:fig5}C we simulate the network in several trials where the number of external inputs change. In every trial, the external inputs to each of the excitatory layers is a number randomly drawn between the layer specific and the layer independent inputs reported in Table \ref{table:externalinput}. The inhibitory layers have their number of external inputs randomly decided between the already chosen number of their respective excitatory layer and the number calculated by

@@

To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test (KS-statistics). To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}. The null hypothesis that two sample sets were drawn from the same distribution is accepted, for an error $\alpha = 0.01$, if the KS-statistic $\leq 1.63$.
\section{Results}\label{results}

@@

\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.001]{fig2.pdf}
 \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article). In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz). (A) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells). Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer). (B–D) Statistics based on samples of 1000 neurons per layer recorded for 60 s. (B) Single-cell firing rate boxplot. Triangles indicate population mean firing rates, crosses show outliers. (C) Irregularity of single-cell firing rates. (D) Synchrony of spiking activity.}\label{fig:fig2}
\end{figure}

In order to compare the Brian 2 with the NEST implementations of the Potjans-Diesmann model, we used the available pyNEST code \cite{potjans2014pyNEST} to run NEST simulations of the model with the same parameters given in Methods. The comparisons were made in terms of the cumulative distributions of the CVs of ISIs and the firing rates of the eight cell populations in the network (plots not shown here but as in Figs.~\ref{fig:fig3} and~\ref{fig:fig4} below). The p-value (two-sample Kolmogorov-Smirnov test) was higher than 0.6 for all comparisons meaning that no significant difference was found between NEST and Brian 2 simulations.

@@

When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences for an error $\alpha = 0.01$. The major discrepancy was found for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.

@@

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.005]{fig5.pdf}
 \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig. 7 of the original article). Color codes and sample cell sizes as in Fig. \rot{\ref{fig:fig2}}. (A1--A2) DC input. (B1--B2) Layer independent input. (C1--C4) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods). }\label{fig:fig5}
\end{figure}

In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs. These are shown in Fig.~\ref{fig:fig5}A, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}B, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs. In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
In Fig.~\ref{fig:fig5}C we present the population firing rates for 100 trials with the rule explained in the methods section. This latter experiment presented an excellent agreement with the histograms observed in Fig. 7 of the original article.

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.1]{fig6.pdf}
 \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig. 8 of the original article). (A) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$. (B) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$. Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate. (C) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (A)).}\label{fig:fig6}
\end{figure}

@@

As described in the results section of the original article, when thalamic input is delivered to the network the output is rather stereotypical. This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian2 to transient thalamic input. By comparing histograms in Fig.~\ref{fig:fig7}\textbf{B} with the ones in Fig. 10 \textbf{B} of the original article one can see the similarities.

\begin{figure}[!ht]
 \centering
% * <aroquesilva@gmail.com> 2018-02-06T01:02:54.595Z:
%
% ^.
 \includegraphics[scale=0.01]{fig7.pdf}
 \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.10 from the original article). Note that the time axis is centralized at the beginning of the thalamic input, which lasts for 10 ms. (A) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section. (B) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
\end{figure}

@@

\section{Acknowledgements}\label{acknowledgements}
The authors would like to thank Prof. Markus Diesmann, Dr. Maximilian Schmidt and Dr. Sacha van Albada for valuable comments on the manuscript. This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to RS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC. ACR is partially supported by the CNPq fellowship grant 306251/2014-0. RP and ACR are also part of the IRTG 1740 / TRP 2015/50122-0, funded by DFG / FAPESP. CCC, and CR are supported by a CAPES PhD scholarship.

>>>>

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[â†’] The cell-type specific cortical microcircuit: relating structure and activity in a full-scale spiking network model, T.C. Potjans,  M. Diesmann, Cerebral Cortex, 24 (3): 785-806, 2014. DOI: 10.1093/cercor/bhs358
  \end{itemize}\par
\end{flushleft}
\end{rebox}

@@

Additionally to these inputs, in Fig.~\ref{fig:fig5}\textbf{C} we simulate the network in several trials where the number of external inputs change. In every trial, the external inputs to each of the excitatory layers is a number randomly drawn between the layer specific and the layer independent inputs reported in Table \ref{table:externalinput}. The inhibitory layers have their number of external inputs randomly decided between the already chosen number of their respective excitatory layer and the number calculated by

@@

To compare the distributions of firing rates and CVs obtained from simulations in Brian 2 and NEST we use the Kolmogorov-Smirnov statistical test (KS-statistics). To apply this non-parametric test, cumulated histograms were constructed with bins chosen by the Doane method \cite{doane1976}.

\section{Results}\label{results}

@@

\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.001]{fig2.pdf}
 \caption{Spontaneous cell-type specific activity (to be compared with Fig.6 from the original article). In this simulation, $g=4$ and background rate is layer-specific ($8$ Hz). (\textbf{A}) Raster plot of spiking activity for 0.4 s of all cell layers (from top to bottom; dark color: Excitatory cells, light color: Inhibitory cells). Number of displayed spike trains corresponds to 2.5\% of the total number of neurons (preserving relative number of cells per layer). (\textbf{B--€“D}) Statistics based on samples of 1000 neurons per layer recorded for 60 s. (\textbf{B}) Single-cell firing rate boxplot. Triangles indicate population mean firing rates, crosses show outliers. (\textbf{C}) Irregularity of single-cell firing rates. (\textbf{D}) Synchrony of spiking activity.}\label{fig:fig2}
\end{figure}

In order to compare the Brian 2 with the NEST implementations of the Potjans-Diesmann model, we used the available pyNEST code \cite{potjans2014pyNEST} to run NEST simulations of the model with the same parameters given in Methods. The comparisons were made in terms of the cumulative distributions of the CVs of ISIs and the firing rates of the eight cell populations in the network (plots not shown here but as in Figs.~\ref{fig:fig3} and~\ref{fig:fig4} below). The two-sample Kolmogorov-Smirnov test pointed that no difference was found between NEST and Brian 2 simulations.

@@

When comparing the network constructed in Brian 2 with equation~(\ref{Eq:Ksyn_alt}) and the one constructed in NEST, all layer-specific comparisons yielded KS-statistics lower than 1.14, indicating no significant differences (for an error $\alpha = 0.01$) between the firing rate and CV distributions. However, when comparing the mean firing rates and their standard deviations, in general, slightly differences were found being major for L5e cells, which in the replicated Brian 2 version using equation~(\ref{Eq:Ksyn_alt}) have mean firing rate 12.2 $\pm$ 6.1 Hz, while using equation~(\ref{Eq:Ksyn}) their mean firing rate is 7.8 $\pm$ 4.8 Hz and in the original version their mean firing rate is 7.8 $\pm$ 5.1 Hz.

@@

\begin{figure}[!htb]
 \centering
 \includegraphics[scale=0.005]{fig5.pdf}
 \caption{Raster plot of spiking activity over a 0.4 s period and average firing rate for different types of external input (to be compared with Fig. 7 of the original article). Color codes and sample cell sizes as in Fig. \rot{\ref{fig:fig2}}. (\textbf{A1--A2}) DC input. (\textbf{B1--B2}) Layer independent input. (\textbf{C1--C4}) Histograms containing the population firing rates of 100 trials where the external inputs were drawn with a specific rule (see details in methods). }\label{fig:fig5}
\end{figure}

In agreement with the results obtained in the reference article, the activity features of the Brian 2 implementation are also robust to changes in the external inputs. These are shown in Fig.~\ref{fig:fig5}\textbf{A}, in which the Poissonian inputs are replaced by constant DC currents, and in Fig.~\ref{fig:fig5}\textbf{B}, in which the layer-dependent Poissonian inputs are replaced by layer-independent inputs. In the latter case, the absence of activity in L6e resulting from the layer-independent inputs indicates the importance of realistic input structure to yield plausible activity in all layers.
In Fig.~\ref{fig:fig5}\textbf{C} we present the population firing rates for 100 trials with the rule explained in the methods section. This latter experiment presented an excellent agreement with the histograms observed in Fig. 7 of the original article.

@@

\begin{figure}[!ht]
 \centering
 \includegraphics[scale=0.1]{fig6.pdf}
 \caption{Network activity dependence on background rate and relative inhibitory weight (to be compared with Fig. 8 of the original article). (\textbf{A}) Mean population firing rates of excitatory neurons in layers 2/3 (dashed line), 4 (triangles), 5 (stars) and 6 (squares) as a function of the background rate for fixed $g=4$. (\textbf{B}) AIness\% (see Methods) as a function of the background rate and the relative inhibitory synaptic strength $g$. Labeled dashed contour lines indicate areas where 25\%, 50\% and 75\% of all populations fire in an asynchronous and irregular mode at low rate. (\textbf{C}) Mean population firing rates of excitatory neurons as a function of the relative inhibitory synaptic strength $g$ for fixed background rate $8$ Hz (markers as in (\textbf{A})).}\label{fig:fig6}
\end{figure}

@@

As described in the results section of the original article, when thalamic input is delivered to the network the output is rather stereotypical. This is confirmed by our results presented in Fig.~\ref{fig:fig7}, which shows the response of the network replicated in Brian 2 to transient thalamic input. By comparing histograms in Fig.~\ref{fig:fig7}\textbf{B} with the ones in Fig. 10\textbf{B} we think that our model replicates the original article.

\begin{figure}[!ht]
 \centering
% * <aroquesilva@gmail.com> 2018-02-06T01:02:54.595Z:
%
% ^.
 \includegraphics[scale=0.01]{fig7.pdf}
 \caption{Response to transient thalamic input (to be compared with panels \textbf{A} and \textbf{B} of Fig.~10 from the original article). Note that the time axis is centralized at the beginning of the thalamic input, which lasts for 10 ms. (\textbf{A}) Sample raster plot of the network while receiving transient thalamic input as described in the Methods section. (\textbf{B}) Cell-type specific population spike-counts averaged over 100 different applications of thalamic stimuli.}\label{fig:fig7}
\end{figure}

@@

\section{Acknowledgements}\label{acknowledgements}
The authors would like to thank Prof. Markus Diesmann, Dr. Maximilian Schmidt and Dr. Sacha van Albada for valuable comments on the manuscript. This research was supported by the S\~{a}o Paulo State Research Foundation (FAPESP) grant No.2013/07699-0 for the Research, Innovation and Dissemination Center for Neuromathematics, FAPESP grant 2016/03855-5 to NLK, FAPESP scholarship 2017/07688-9 to ROS, FAPESP scholarship 2013/25667-8 to RP and FAPESP scholarship 2017/05874-0 to VC. ACR is partially supported by the CNPq fellowship grant 306251/2014-0. RP and ACR are also part of the IRTG 1740 / TRP 2015/50122-0, funded by DFG / FAPESP. CCC, and CR are supported by a CAPES PhD scholarship.
